---
title: 'PROJECT #3 for Applied Nonparametric Econometrics'
author: "Jia Ru"
date: '2016-4-18'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, cache = TRUE, results='hold', 
  message = FALSE, warning = FALSE, error = FALSE)
```

# 1. Do Monte Carlo simulations to compare the performances of the local linear and local constant estimations 

The nonparametric regression function is
$$ Y_t = m(X_t) + u_t, \quad 1 \leq t \leq T $$
where **$X_t$ is stationary**.Choose different sample sizes, different kernels, different bandwidths, and different bandwidth selection methods; Use MADE to measure the performance of estimation. Boxplot the distribution of MADE for different settings.

ANS:

Here I set DGP as: $X_t$ is IID follows Uniform distributon on $[0,3]$, $m(x)=e^x$, and $u_t$ is IID follows Standard Normal Distribution. I refer to *4.7 Computer Code in Lecture note, page 88* to define functions in order to get the local-constant and local-linear estimator, with a few modifications, that is: 1. set some default arugments: grid points is exactly sample point x, and bandwidth is 0.02 by default, kernel is normal by default for convenience; 2. modify the local.linear function by directly return the fitted value, ie. Intercept instead of the n-by-2 matrix beta. I also define functions to generate data under different settings and to calculate MADE.

```{r 1.Define:ll,lc,DGP,MADE}
rm(list=ls())
EpanKer<-function(x){0.75*(1-x^2)*(abs(x)<=1)}

# ####################################################
# Define function of local constant estimator
# ####################################################

local.constant=function(y,x,z=x,h=0.02,ker="norm"){
  nz <- length(z)
  nx <- length(x)
  x0 = rep(1,nx*nz)
  dim(x0) = c(nx,nz)
  x1 = t(x0)
  x0 = x*x0
  x1 = z*x1
  x0 = x0-t(x1)
  if(ker=="Epan"){x1=EpanKer(x0/h)}
  if(ker=="norm"){x1=dnorm(x0/h)}
  x2 = y*x1
  f1 = apply(x1,2,mean)
  f2 = apply(x2,2,mean)
  f3 = f2/f1
  
  return(f3)  # return num[1:n]
} 

# ####################################################
# Define the local linear estimator
# ####################################################

local.linear<-function(y,x,z=x,h=0.02,ker="norm"){
  nz <- length(z)
  ny <- length(y)
  beta <- rep(0,nz*2)
  dim(beta) <- c(nz,2)
  for(k in 1:nz){
    x0 = x-z[k]
    if(ker=="Epan"){w0 <- EpanKer(x0/h)}
    if(ker=="norm"){w0 <- dnorm(x0/h)}
    beta[k,] <- glm(y~x0,weight=w0)$coeff
  }
  return(beta[,1])  
}

# ####################################################
# Define function DGP: Data Generating Process
# ####################################################

DGP <- function(n,stationary=TRUE){
  if (stationary==TRUE) {
    x <- runif(n, min=0, max=3)
    mx <- exp(x)
    y <- mx + rnorm(n)
  } else if (stationary==FALSE){
    x <- runif(1, min = 0, max=3)
    for (i in 2:n) {x[i] <- x[i-1]+rnorm(1)}
    mx <- exp(x)
    y <- mx
  }
  return(rbind(x,y,mx))
}

# ####################################################
# Define function MADE, MADE_set
# ####################################################

# Generate One MADE
MADE <- function(n,ker,h=0.02,stationary=T,method) {
  sample <- DGP(n,stationary)
  y <- sample["y",]
  x <- sample["x",]
  m <- sample["mx",]
  m_hat <- 
    if (method=="ll") {
      local.linear(y,x,ker=ker,h=h)
    } else if (method=="lc") {
      local.constant(y,x,ker=ker,h=h)
    }
  MADE <- 1/n*sum(abs(m-m_hat))
  return(MADE)
}

# Generate a set of n_sim MADEs
library(plyr)
MADE_set <- function(n_sim,
                     n,ker,h=0.02,stationary=T,method)
{
  MADE_set <- laply(
    .data = as.list(1:n_sim),
    .fun = function(foo) MADE(n,ker,h,stationary,method),
    .progress = "text"
  )
  return(MADE_set)
}

# test:
# MADE_set(n_sim=10,
#          n=20,ker="Epan",h=0.02,stationary = T,method="ll")
```

Now use `MADE_set()` function to do Monte Carlo Simulations:

#### (1) different sample sizes

```{r 1. MC_n, eval=FALSE}
n_sim = 100
n_set <- list(20,50,100,200,500)

# 1. local linear
MADE_n_ll <- laply(
  .data = n_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=x,ker="Epan",h=0.02,
                  stationary = T,method="ll")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_n_ll) <- paste0("n=",n_set)

boxplot(MADE_n_ll,use.cols=F,outline=FALSE,
        xlab="sample size",ylab="MADE")
title("1-1-1 MADE of different sample sizes, Local Linear")

# 2. local constant
MADE_n_lc <- laply(
  .data = n_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=x,ker="Epan",h=0.02,
                  stationary = T,method="lc")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_n_lc) <- paste0("n=",n_set)
boxplot(MADE_n_lc,use.cols=F,outline=FALSE,
        xlab="sample size",ylab="MADE")
title("1-1-2 MADE of different sample sizes, Local Constant")

```
![](1-1-1.pdf)
![](1-1-2.pdf)

Here I try 5 different sample sizes: 20,50,100,200,500. From the figure we can see, as sample size increase, both mean and standard variance of MADE decrease, it is the same for local constant method and local linear method.

#### (2) different bandwidths

```{r 1.MC_h, eval=FALSE}

# 1. local linear
n_sim = 20
h_set <- as.list(seq(0.1,1.5,by=0.1))

MADE_h_ll <- laply(
  .data = h_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=50,ker="Epan",h=x,
                  stationary = T,method="ll")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_h_ll) <- h_set

boxplot(MADE_h_ll,use.cols=F,outline=FALSE,
        xlab="bandwidth",ylab="MADE")
title("1-2-1 MADE of different bandwidth, Local Linear")

# 2. local constant
n_sim = 20
h_set <- as.list(seq(0.01,0.9,by=0.04))

MADE_h_lc <- laply(
  .data = h_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=50,ker="Epan",h=x,
                  stationary = T,method="lc")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_h_lc) <- h_set

boxplot(MADE_h_lc,use.cols=F,outline=FALSE,
        xlab="bandwidth",ylab="MADE")
title("1-2-2 MADE of different bandwidth, Local Constant")
```

![](1-2-1.pdf)
![](1-2-2.pdf)

from the two figure we can see, the best bandwidth for local constant method is approximately 0.2, the best bandwidth for local linear method is approximately 0.4.

