---
title: 'PROJECT #3 for Applied Nonparametric Econometrics'
author: "Jia Ru"
date: '2016-4-18'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, cache = TRUE, results='hold', 
  message = FALSE, warning = FALSE, error = FALSE, include = FALSE)
```

# 1. Do Monte Carlo simulations to compare the performances of the local linear and local constant estimations 

The nonparametric regression function is
$$ Y_t = m(X_t) + u_t, \quad 1 \leq t \leq T $$
where **$X_t$ is stationary**.Choose different sample sizes, different kernels, different bandwidths, and different bandwidth selection methods; Use MADE to measure the performance of estimation. Boxplot the distribution of MADE for different settings.

ANS:

Here I set DGP as: $X_t$ is IID follows Uniform distributon on $[0,3]$, $m(x)=e^x$, and $u_t$ is IID follows Standard Normal Distribution. I refer to *4.7 Computer Code in Lecture note, page 88* to define functions in order to get the local-constant and local-linear estimator, with a few modifications, that is: 1. set some default arugments: grid points is exactly sample point x, and bandwidth is 0.02 by default, kernel is normal by default for convenience; 2. modify the local.linear function by directly return the fitted value, ie. Intercept instead of the n-by-2 matrix beta. I also define functions to generate data under different settings and to calculate MADE.

```{r 1-0,eval=FALSE,include=T}
rm(list=ls())
EpanKer<-function(x){0.75*(1-x^2)*(abs(x)<=1)}

# ####################################################
# Define function of local constant estimator
# ####################################################

local.constant=function(y,x,z=x,h=0.02,ker="norm"){
  nz <- length(z)
  nx <- length(x)
  x0 = rep(1,nx*nz)
  dim(x0) = c(nx,nz)
  x1 = t(x0)
  x0 = x*x0
  x1 = z*x1
  x0 = x0-t(x1)
  if(ker=="Epan"){x1=EpanKer(x0/h)}
  if(ker=="norm"){x1=dnorm(x0/h)}
  x2 = y*x1
  f1 = apply(x1,2,mean)
  f2 = apply(x2,2,mean)
  f3 = f2/f1
  
  return(f3)  # return num[1:n]
} 

# ####################################################
# Define the local linear estimator
# ####################################################

local.linear<-function(y,x,z=x,h=0.02,ker="norm"){
  nz <- length(z)
  ny <- length(y)
  beta <- rep(0,nz*2)
  dim(beta) <- c(nz,2)
  for(k in 1:nz){
    x0 = x-z[k]
    if(ker=="Epan"){w0 <- EpanKer(x0/h)}
    if(ker=="norm"){w0 <- dnorm(x0/h)}
    beta[k,] <- glm(y~x0,weight=w0)$coeff
  }
  return(beta[,1])  
}

# ####################################################
# Define function DGP: Data Generating Process
# ####################################################

DGP <- function(n,stationary=TRUE){
  if (stationary==TRUE) {
    x <- runif(n, min=0, max=3)
    mx <- exp(x)
    y <- mx + rnorm(n)
  } else if (stationary==FALSE){
    x <- runif(1, min = 0, max=3)
    for (i in 2:n) {x[i] <- 0.1*x[i-1]+rnorm(1)}
    mx <- x^2
    y <- mx
  }
  return(rbind(x,y,mx))
}

# ####################################################
# Define function MADE, MADE_set
# ####################################################

# Generate One MADE
MADE <- function(n,ker,h=0.02,stationary=T,method) {
  sample <- DGP(n,stationary)
  y <- sample["y",]
  x <- sample["x",]
  m <- sample["mx",]
  m_hat <- 
    if (method=="ll") {
      local.linear(y,x,ker=ker,h=h)
    } else if (method=="lc") {
      local.constant(y,x,ker=ker,h=h)
    }
  MADE <- 1/n*sum(abs(m-m_hat))
  return(MADE)
}

# Generate a set of n_sim MADEs
library(plyr)
MADE_set <- function(n_sim,
                     n,ker,h=0.02,stationary=T,method)
{
  MADE_set <- laply(
    .data = as.list(1:n_sim),
    .fun = function(foo) MADE(n,ker,h,stationary,method),
    .progress = "text"
  )
  return(MADE_set)
}

# test:
# MADE_set(n_sim=10,
#          n=20,ker="Epan",h=0.02,stationary = T,method="ll")
```

Now use `MADE_set()` function to do Monte Carlo Simulations: (Code hide)

#### 1-1 different sample sizes

```{r 1-1, eval=FALSE}
n_sim = 100
n_set <- list(20,50,100,200,500)

# 1. local linear
MADE_n_ll <- laply(
  .data = n_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=x,ker="Epan",h=0.02,
                  stationary = T,method="ll")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_n_ll) <- paste0("n=",n_set)

boxplot(MADE_n_ll,use.cols=F,outline=FALSE,
        xlab="sample size",ylab="MADE")
title("1-1-1 MADE of different sample sizes, Local Linear")

# 2. local constant
MADE_n_lc <- laply(
  .data = n_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=x,ker="Epan",h=0.02,
                  stationary = T,method="lc")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_n_lc) <- paste0("n=",n_set)
boxplot(MADE_n_lc,use.cols=F,outline=FALSE,
        xlab="sample size",ylab="MADE")
title("1-1-2 MADE of different sample sizes, Local Constant")

```

![](1-1-1.pdf)
![](1-1-2.pdf)

Here I try 5 different sample sizes: 20,50,100,200,500. From the figure we can see, as sample size increase, both mean and standard variance of MADE decrease, it is the same for local constant method and local linear method.

#### 1-2 different bandwidths

```{r 1-2, eval=FALSE}

# 1. local linear
n_sim = 20
h_set <- as.list(seq(0.1,1.5,by=0.1))

MADE_h_ll <- laply(
  .data = h_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=50,ker="Epan",h=x,
                  stationary = T,method="ll")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_h_ll) <- h_set

boxplot(MADE_h_ll,use.cols=F,outline=FALSE,
        xlab="bandwidth",ylab="MADE")
title("1-2-1 MADE of different bandwidth, Local Linear")

# 2. local constant
n_sim = 20
h_set <- as.list(seq(0.01,0.9,by=0.04))

MADE_h_lc <- laply(
  .data = h_set,
  .fun = function(x) {
    yy <- MADE_set(n_sim=n_sim,n=50,ker="Epan",h=x,
                  stationary = T,method="lc")
    return(yy)
  },
  .progress = "text"
)
rownames(MADE_h_lc) <- h_set

boxplot(MADE_h_lc,use.cols=F,outline=FALSE,
        xlab="bandwidth",ylab="MADE")
title("1-2-2 MADE of different bandwidth, Local Constant")
```

![](1-2-1.pdf)
![](1-2-2.pdf)

from the two figure we can see, the best bandwidth for local constant method is approximately 0.2, the best bandwidth for local linear method is approximately 0.4.

#### 1-3 different kernels

```{r 1-3, eval=FALSE}
MADE_Epan_ll <- MADE_set(n_sim=30,n=50,ker="Epan",h=0.4,
                  stationary = T,method="ll")
MADE_Norm_ll <- MADE_set(n_sim=30,n=50,ker="norm",h=0.4,
                  stationary = T,method="ll")
MADE_Epan_lc <- MADE_set(n_sim=30,n=50,ker="Epan",h=0.2,
                  stationary = T,method="lc")
MADE_Norm_lc <- MADE_set(n_sim=30,n=50,ker="norm",h=0.2,
                  stationary = T,method="lc")
MADE_ker <- rbind(MADE_Epan_ll,MADE_Norm_ll,MADE_Epan_lc,MADE_Epan_lc)
rownames(MADE_ker) <- c("Epan_ll","Norm_ll","Epan_lc","Norm_lc")
boxplot(MADE_ker,use.cols=F,outline=FALSE,
        xlab="different kernels & different method",ylab="MADE")
title("1-3 different kernels & different method")

```

![](1-3.pdf)

from the figure we can see, Epan kernel with local linear method performs best. But local constant method has smaller variance. Here I use different bandwidth for the two method which is "best" in terms of MADE, obtained in last question.


# 2. Re-consider Question 1, where $X_t$ is nonstationary 

All the operation is the same as in Question 1, except that the data generating process is:
$$ X_t = 0.1 X_{t-1} + v_t $$
where $v_t$ is iid standard normal. And $Y_t$ is 
$$ Y_t = X_t^2 + u_t $$

(Codes hide)

#### 2-1 different sample sizes
![](2-1-1.pdf)
![](2-1-2.pdf)

It is totally different from the stationary case that as sample size increase, the MADE also increase. This is because of serial correlation of $X_t$. The local linear method seems to have smaller MADE for all sample sizes.

#### 2-2 different bandwidths

![](2-2-1.pdf)
![](2-2-2.pdf)

from the two figure we can see, the optimal bandwidth seem to be as small as possible.(I have tried even smaller bandwidth but it's almost the same pattern). 

#### 2-3 different kernels

![](2-3.pdf)

from the figure we can see, Epan kernel with local linear method has smallest MADE.


# 3. Empirical Analysis: The nonlinearity of spread of interest rates 

I download 1-Year Treasury Constant Maturity Rate from <https://research.stlouisfed.org/fred2/series/WGS1YR#> and 3-Year Treasury Constant Maturity Rate from <https://research.stlouisfed.org/fred2/series/WGS3YR#>. 

First of all, take an overlook at the data. Plot $r_{1t}$, $r_{3t}$, $s_t$ and $s^\star_t$. The data looks like this:

```{r 3-0, include=T, eval=T, echo=F,results="hold"}

rm(list=ls())
require(readxl)
require(ggplot2)

WGS1YR <- read_excel("Q3_data/WGS1YR.xls", 
                     col_names = c("date","r1"), skip=11)
WGS3YR <- read_excel("Q3_data/WGS3YR.xls", 
                     col_names = c("date","r3"), skip=11)
WGS <- merge(WGS1YR, WGS3YR, by="date", all=TRUE)
WGS <- mutate(WGS, logr1=log(r1), logr3=log(r3), s=logr3-logr1)
WGS$s_star <- c(NA,WGS$s[-1]-WGS$s[-nrow(WGS)])
WGS$time <- 1:nrow(WGS)

p1 <- ggplot(data = WGS, mapping = aes(x=date)) +
  geom_line(aes(y=r1), col="red", alpha=0.8) +
  geom_line(aes(y=r3), col="orange", alpha=0.8) +
  annotate("text", label = "r1", x = WGS$date[1], y = 5, size = 5, colour = "red")+
  annotate("text", label = "r3", x = WGS$date[1], y = 6, size = 5, colour = "orange")+
  ylab("r1 & r3") +
  ggtitle("1-Year & 3-Year Treasury Constant Maturity Rate, loged")
p1

p2 <- ggplot(data=WGS, aes(x=date,y=s)) + 
  geom_line(col="blue") +
  ggtitle("log interest rate spread s")
p2

p3 <- ggplot(data=WGS, aes(x=date,y=s_star)) + 
  geom_line(col="darkolivegreen") +
  #scale_y_continuous(limits = c(-0.2,0.2)) +
  ggtitle("first-order difference of log interest rate spread")
p3

```

To study if $\{s_t\}$ is nonlinear, I try different nonparametric regression model and do analysis of variance to test whether local linear model is suitable. Since the $\{s_t\}$ is time series, I first see the ACF and PACF plot of the series. From the ACF plot we can see, the auto-correlation is significant positive and very large for all lag less than about 100 while the 1st-order lag is almost 0.9, so I construct the regression model using only 1st-order lag as explainatory variable:
$$ s_t = m(s_{t-1}) + \epsilon_t $$
Here I use `dyn` and `loess` package in R to do the regression. Use `anova()` to test Hypothesis.

```{r 3-1, include=T,eval=T,results="hold"}
require(np)
require(loess)
require(locfit)
library(dyn)

#####################  1. s   ######################
st <- ts(WGS$s)

# ACF and PACF
par(mfrow=c(2,2))
acf(st, lag.max = nrow(WGS), main="ACF of s")
pacf(st, lag.max = nrow(WGS), main="PACF of s")
acf(st, main="ACF of s")
pacf(st, main="PACF of s")

# noneparametric regression
np_l <- dyn$loess(st ~ lag(st, -1),degree=1)  # local linear
np_nl <- dyn$loess(st ~ lag(st, -1),degree=2)  # local polynomial
fitted_l <- fitted(np_l)
fitted_nl <- fitted(np_nl)

# visialization
ggplot(data=NULL, aes(x=st[-1],y=st[-length(st)])) + 
  geom_rug(alpha=0.3) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_line(aes(y=fitted_l), col="green",alpha=0.7) +
  geom_line(aes(y=fitted_nl), col="red", alpha=0.7) +
  xlab("lag st") +
  ylab("st") +
  ggtitle("3 regression of s[t] on s[t-1]")

# Hypothesis Testing
anova(np_l,np_nl)[-1]
```

Results:

the linear and nonlinear approach gives almost the same result and very close to the sample point except some outliers. The Hypothesis Testing result show that we have to **reject** the null hypothesis that ${s_t}$ is linear under the significance level 0.05.

For series $s^\star_t$ do the same thing:

```{r 3-2, include=T,eval=T,results="hold"}
#####################  2. s_star   ######################
st <- ts(WGS$s_star)[-1]

# ACF and PACF
par(mfrow=c(2,2))
acf(st, lag.max = nrow(WGS), main="ACF of s_star")
pacf(st, lag.max = nrow(WGS), main="PACF of s_star")
acf(st, main="ACF of s_star")
pacf(st, main="PACF of s_star")

# nonparametrci regression
lm <- lm(st ~ lag(st, -1))
np_l <- dyn$loess(st ~ lag(st, -1),degree=1)
np_nl <- dyn$loess(st ~ lag(st, -1),degree=2)
fitted_l <- fitted(np_l)
fitted_nl <- fitted(np_nl)
fitted_lm <- fitted(lm)

# visalization
ggplot(data=NULL, aes(x=st[-1],y=st[-length(st)])) + 
  geom_rug(alpha=0.3) +
  geom_point(alpha=0.2) +
  geom_smooth(method="lm") +
  geom_line(aes(y=fitted_l[-1]), col="green",alpha=0.7) +
  geom_line(aes(y=fitted_nl[-1]), col="red", alpha=0.7) +
  xlab("lag s_star") +
  ylab("s_star") +
  ggtitle("3 regression of s_star[t] on s_star[t-1]")

# Hypothesis Testing
anova(np_l,np_nl)[-1]

```

Results:

the linear and nonlinear approach gives exactly the same result, but from the scatter plot we can see the joint distribution of $s^{\star}_t$ and $s^{\star}_{t-1}$ is clustered at point $(0,0)$, so the both the nonparametric models and paramethric linear model does not fit well. The Hypothesis Testing result show that we **reject** the null hypothesis that ${s^\star_t}$ is linear under the significance level 0.05.



# 4. Consider three real data sets for the U.S. Treasury bill (Secondary Market Rate): the **daily, weekly and monthly** 3-month Treasury bill.

First let's see what the datas look like:

```{r 4-0}

rm(list = ls())

require(ggplot2)
require(lubridate)

daily <- read.csv(file="Project_3_DATA/DTB3.csv", 
                  header = T, na.strings = ".", stringsAsFactors = F)
weekly <- read.csv(file="Project_3_DATA/WTB3MS.csv", 
                  header = T, na.strings = ".", stringsAsFactors = F)
monthly <- read.csv(file="Project_3_DATA/TB4WK.csv", 
                  header = T, na.strings = ".", stringsAsFactors = F)
daily$DATE <- as.Date(daily$DATE)
monthly$DATE <- as.Date(monthly$DATE)
weekly$DATE <- as.Date(weekly$DATE)

m <- ts(monthly$VALUE, 
        start = c(year(monthly$DATE[1]), month(monthly$DATE[1])), 
        frequency = 12)
w <- ts(weekly$VALUE, 
        start = c(year(weekly$DATE[1]), week(weekly$DATE[1])),
        frequency = 52)
d <- ts(daily$VALUE, 
        start = c(year(daily$DATE[1]), month(daily$DATE[1]), day(daily$DATE[1])), 
        frequency = 365)

par(mfrow=c(3,1))
ts.plot(m, main="monthly")
ts.plot(w, main="weekly")
ts.plot(d, main="daily")
# 
# qplot(x=DATE, y=VALUE, data = daily, geom="line", main = "daily")
# qplot(x=DATE, y=VALUE, data = monthly, geom="line", main = "monthly")
# qplot(x=DATE, y=VALUE, data = weekly, geom="line", main = "weekly")
```

![](4-0.pdf)

### 4-1. Apply the nonparametric regression estimation methods to estimate the drift and diffusion functions for each series and higher moments such as skewedness and kurtosis.

I will use the function `ksdrift` and `ksdiff`  in R package `sde` , whose description in the document is "Implementation of simple Nadaraya-Watson nonparametric estimation of drift and diffusion coefficient", the default bandwidth is calculated using Scott’s rule (i.e., $bandwidth = n^{-1/5}*sd(x)$). The estimation code and result is plot as follows:

```{r 4-1}

require(plyr)
require(sde)

drift_diffussion <- function(x) {
  drift <- ksdrift(x)
  diffusion <- ksdiff(x)

  plot(drift, type="l", main="drift")
  plot(diffusion, type="l", main="diffusion")
}

d_new <- na.locf(d)   # replacing each NA with the most recent non-NA prior to it.
par(mfrow=c(3,2))
l_ply(.data=list(m,w,d_new), 
      .fun=drift_diffussion,
      .progress = "text"
      )
```

(First row is **monthly** data, second is **weekly**, the bottom line is **daily**.)

![](4-1.pdf)

Next we estimate the skewness and kurtoiss of the 3 data: Here I use `skewness` and `kurtosis` function in `moments` package:

```{r 4-1_, include=T, results="markup"}

require(moments)

skewness(m); kurtosis(m)
skewness(w);kurtosis(w)
skewness(d, na.rm = T);kurtosis(d, na.rm = T)

```

### 4-2. Any conclusions and comments on three drift and diffusion functions? Also, think about your conclusions on whether a jump diffusion model is appropriate for the data; see Johannes (2004).

I use R command `BNSjumptest` in package `highfreqency`to test the presence of jumps. It is Barndorff- Nielsen and Shephard (2006) tests for the presence of jumps in the price series,which examines the presence of jumps in highfrequency price series. It is based on theory of Barndorff- Nielsen and Shephard (BNS). The null hypothesis is no jumps.

The test result is that we **reject** the null hypothesis, i.e we consider there exist jumps in the series of intrest rates.

```{r 4-2,include=T, results="markup"}
require(highfrequency)

BNSjumptest(m)
BNSjumptest(w)
BNSjumptest(d_new)


```

<!--
### 4-3. Compare your results with results in Ait-Sahalia (1996), Jiang and Knight (1997), and Stanton (1997).
-->
